A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study

### 2022-08-10 02_paper/02_study/02_fitbit paper/fitbit_paper.Rmd
update plots

needs(tidyverse,
      lubridate,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable)

@@

@@

# Discussion

>>>>

needs(tidyverse,
      lubridate,
      janitor,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable)

@@

### FitBit

All participants were given a FitBit Smart Watch Charge 4 to wear during the experiment.

## Heart Rate

The heart rate of each subject were measured before, during and after the experiment.

XXX Data and data wrangling need to be imported XXX 
```{r}
# Initial Preparations
## Run following Code if and only if new data is added to the initial data.

source(file = "data.wrangling.R")

# Plots

df_heart_all <-
  read_rds(file = "final_data/heart_all.rds") |> 
  mutate(status = factor(status,
                         levels = c("Expert","Novice"),
                         labels = c("Expert","Novice")
                         )
         )

df_heart_min <-
  read_rds(file = "final_data/heart_min.rds")

df_heart_aggr <- 
  read_rds(file = "final_data/heart_aggr.rds") |> 
  mutate(status = factor(status,
                         levels = c("Novice","Expert"),
                         labels = c("Novice","Expert")
                         )
         )

plot_flow <-
ggplot(data = df_heart_all,
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`,
                     group = ID)
       ) +
  geom_line(mapping = aes(color = status),
            size = 0.3,
            alpha = 1
            ) +
  xlim(0, min(df_heart_min$min)) +
  scale_color_viridis_d(option = "A",
                       end = 0.8)  +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.text = element_text(size = 16)) +
  facet_grid(cols = vars(Type),
             scales = "free_x")

plot_flow_leg <-
ggplot(data = df_heart_all,
       mapping = aes(x = Time,
                     y = `Heart Rate`
                     )
       ) +
  geom_smooth(mapping = aes(color = status),
              method = "loess") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
 # scale_linetype_manual(values = c(1, 6)) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.title = element_blank())

legend <- cowplot::get_legend(plot_flow_leg)

plot_loess <-
ggplot(data = filter(df_heart_all),
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`
                     )
       ) +
  stat_smooth(geom = "line",
              mapping = aes(color = status),
              method = "loess",
              se = FALSE) +
  xlim(0,min(df_heart_min$min)) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  facet_grid(cols = vars(Type),
             scales = "free_x",
             ) +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

plot_box <- 
ggplot(data = df_heart_aggr,
       mapping = aes(x = Type,
                     y = mean,
                     color = status
                     )
       ) +
  geom_boxplot(color = "darkgrey",
               fill = "grey") +
  geom_line(data = filter(df_heart_aggr,
                          status == "Novice"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 1,
            alpha = 0.5
            ) +
  geom_line(data = filter(df_heart_aggr,
                          status == "Expert"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 6,
            alpha = 0.5
            ) +
  geom_point(size = 2,
             alpha = 0.5) +
  labs(x = "",
       y = "Average Heart Rate\n(in Beats per Minute)") +
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        strip.text = element_blank(),
       legend.position = "none",
        axis.text.x = element_blank()
        ) +
  aes(fill = Type)

lo_mat <- rbind(c(1,1,1,1,1,1,NA),
                c(2,2,2,2,2,2,4),
                c(3,3,3,3,3,3,NA))

plot_heart <- 
grid.arrange(grobs = list(plot_flow, plot_loess, plot_box, legend),
             layout_matrix = lo_mat)

plot_heart
```

### Steps

Using the FitBit we measured the steps a person walked during the experiment as well. We validated the steps in six randomly selected videos by manual step counting. A step was considered to be the movement of the subject's foot forward, backward, or sideways. The measurement was not as reliable as hoped.

```{r}
df_steps <- read_rds(file = "final_data/steps.rds")
set.seed(5678)
ggplot(data = df_steps,
       mapping = aes(x = type,
                     y = Steps
                     )
       ) +
  geom_boxplot(mapping = aes(color = type),
               fill = "grey",
               alpha = 0.6) +
  geom_jitter(mapping = aes(color = type),
              alpha = 0.9,
              width = 0.3) + 
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  labs(x = "",
       y = "Steps During\nMain Phase") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
        )

### Validation of steps 
# selecting some rows at random with sample_n()
sample_n(df_steps, 6)

# random ID Output:
# Novices ID: 102, 125, 110
# Experts ID: 219, 218, 214

# Results from manual step counting:
# Novices ID: 110 -> 178 steps; 102 -> 219 steps; 125 -> 319 steps
# Experts ID: 214 -> 437 steps; 218 -> 195 steps; 219 -> 530 steps

@@

# Discussion and implication
